Databrary: A digital data library for sharing research video

### 2015-01-14 jcdl-15/paper.tex
Fix inline citations

Data sharing and open science practices are emerging or accepted norms
in the biological, physical and earth sciences with demonstrated
benefits for scientific transparency and accelerated discovery [1-4].
Open research practices are gaining traction in the behavioral and
social sciences [5-10]. However, open data sharing among developmental
researchers is still the exception [11] not common practice. In seeking
to expand sharing of video-based data, Databrary faces challenges
concerning the interplay among institutional research policies, ethical
issues related to obtaining participants’ consent for sharing, and
technical authorization systems. This paper describes the current state
of the Databrary project including the project organization, how content
is acquired, stored, preserved, and accessed, policy considerations, and
how the library has affected and hopes to affect the behavioral science
research community.

@@

The Databrary team recognizes that other data repositories enforce
strict metadata ontologies and that doing so may have benefits in
sub-domains of research when there is community consensus [12]. We will
support standard data coding ontologies among researchers, but only
enforce standardization for a small set of standard tags such as study
date, participant birth date, and sex. We will also encourage
contributors to report race, ethnicity, primary language, language of
dataset, and location of session. The Databrary team chose not to
require strict metadata ontologies for several reasons. We hope to
reduce the pre-deposit curational demands on contributors, encourage the
repurposing of shared data, and foster the rapid adoption of data
sharing. Developing and achieving agreement on metadata ontologies can
take significant amounts of time. Video data are so rich and complex
that in many domains, researchers have not settled on standard
definitions for particular behaviors and may have little current need
for standard tasks, procedures, or terminology. However, we will also
encourage users to re-use tags and terminology by suggesting common or
similar terms, without confining users to these suggestions. User
communities within Databrary may eventually converge on common
conceptual and metadata ontologies based on the most common (and
commonly endorsed) keyword tags, but standardized ontologies are not
necessary for browsing and searching in most of the use cases we
envision.

Our experiences curating and ingesting archival datasets have
highlighted the considerable value of contributors entering raw video
data into Databrary as soon as recordings are acquired. Immediate
uploading reduces the workload on investigators, minimizes the risk of
data loss and corruption, and accelerates the speed with which materials
become openly available. To encourage immediate uploading, Databrary
provides a complete set of controls so that researchers can restrict
access to their data to only their own labs or to other users of their
choosing. Datasets can be shared at a later point when data collection
and ancillary materials are complete, whenever the contributor is
comfortable sharing, or when journals or funders require it. Databrary
has published a Data Sharing Manifesto [13] that explains to researchers
the Databrary philosophy. Standards about when data should be shared are
evolving. Our philosophy is consistent with concepts and practices in
other domains where data sharing is the norm (e.g., www.iedata.org).
Planned enhancements to the Datavyu tool will allow contributors to
organize video files and metadata as part of the analysis process. This
will facilitate the contribution of packaged datasets to Databrary.

@@

For preservation, the original file (if digital) or the preservation
copy will be stored in a long-term preservation repository managed
jointly by the NYU Libraries and the central Information Technology
Services unit. This repository ensures that each content item has a METS
[14] structural metadata file that associates the digital asset with its
metadata. It stores files in two mirrored and geographically distributed
locations, and a third copy on offsite tape; it performs regular fixity
checks; and it provides a format migration capacity, in the event that a
stored format becomes at risk of obsolescence.

@@

Databrary has also begun to have an impact in the policy arena. The
notion that identifiable research data may be shared – under the right
circumstances – is not new to Databrary [15,16]. But, Databrary has
created a set of policies and template documents that will help IRBs to
come to see that seeking permission to share data is merely an extension
of the principle of informed consent. Furthermore, Databrary’s
Authorized Investigator Agreement combines provisions for data
contribution with those guiding data use, part of an overall effort to
reduce barriers to sharing and re-use. Our colleagues in the data
sharing community tell us that this combination represents an innovation
in itself.

>>>>

Data sharing and open science practices are emerging or accepted norms
in the biological, physical and earth sciences with demonstrated
benefits for scientific transparency and accelerated discovery \cite{Curry2011,iedadata,Kaye2009,Overpeck2011}.
Open research practices are gaining traction in the behavioral and
social sciences \cite{Lunshof2014,Adolph2012,King2011,APS2012,Nosek2012i,Nosek2012ii}. However, open data sharing among developmental
researchers is still the exception \cite{MacWhinney2001} not common practice. In seeking
to expand sharing of video-based data, Databrary faces challenges
concerning the interplay among institutional research policies, ethical
issues related to obtaining participants’ consent for sharing, and
technical authorization systems. This paper describes the current state
of the Databrary project including the project organization, how content
is acquired, stored, preserved, and accessed, policy considerations, and
how the library has affected and hopes to affect the behavioral science
research community.

@@

The Databrary team recognizes that other data repositories enforce
strict metadata ontologies and that doing so may have benefits in
sub-domains of research when there is community consensus \cite{cogatlas}. We will
support standard data coding ontologies among researchers, but only
enforce standardization for a small set of standard tags such as study
date, participant birth date, and sex. We will also encourage
contributors to report race, ethnicity, primary language, language of
dataset, and location of session. The Databrary team chose not to
require strict metadata ontologies for several reasons. We hope to
reduce the pre-deposit curational demands on contributors, encourage the
repurposing of shared data, and foster the rapid adoption of data
sharing. Developing and achieving agreement on metadata ontologies can
take significant amounts of time. Video data are so rich and complex
that in many domains, researchers have not settled on standard
definitions for particular behaviors and may have little current need
for standard tasks, procedures, or terminology. However, we will also
encourage users to re-use tags and terminology by suggesting common or
similar terms, without confining users to these suggestions. User
communities within Databrary may eventually converge on common
conceptual and metadata ontologies based on the most common (and
commonly endorsed) keyword tags, but standardized ontologies are not
necessary for browsing and searching in most of the use cases we
envision.

Our experiences curating and ingesting archival datasets have
highlighted the considerable value of contributors entering raw video
data into Databrary as soon as recordings are acquired. Immediate
uploading reduces the workload on investigators, minimizes the risk of
data loss and corruption, and accelerates the speed with which materials
become openly available. To encourage immediate uploading, Databrary
provides a complete set of controls so that researchers can restrict
access to their data to only their own labs or to other users of their
choosing. Datasets can be shared at a later point when data collection
and ancillary materials are complete, whenever the contributor is
comfortable sharing, or when journals or funders require it. Databrary
has published a Data Sharing Manifesto \cite{manifesto} that explains to researchers
the Databrary philosophy. Standards about when data should be shared are
evolving. Our philosophy is consistent with concepts and practices in
other domains where data sharing is the norm (e.g., www.iedata.org).
Planned enhancements to the Datavyu tool will allow contributors to
organize video files and metadata as part of the analysis process. This
will facilitate the contribution of packaged datasets to Databrary.

@@

For preservation, the original file (if digital) or the preservation
copy will be stored in a long-term preservation repository managed
jointly by the NYU Libraries and the central Information Technology
Services unit. This repository ensures that each content item has a METS
\cite{METS} structural metadata file that associates the digital asset with its
metadata. It stores files in two mirrored and geographically distributed
locations, and a third copy on offsite tape; it performs regular fixity
checks; and it provides a format migration capacity, in the event that a
stored format becomes at risk of obsolescence.

@@

Databrary has also begun to have an impact in the policy arena. The
notion that identifiable research data may be shared – under the right
circumstances – is not new to Databrary \cite{HCP,PGP}. But, Databrary has
created a set of policies and template documents that will help IRBs to
come to see that seeking permission to share data is merely an extension
of the principle of informed consent. Furthermore, Databrary’s
Authorized Investigator Agreement combines provisions for data
contribution with those guiding data use, part of an overall effort to
reduce barriers to sharing and re-use. Our colleagues in the data
sharing community tell us that this combination represents an innovation
in itself.
